%%
%% This is file `thesis.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% ubthesis.dtx  (with options: `example')
%% 
%% This is a generated file.
%% 
%% Copyright (C) 2007 by Justin Boffemmyer <jsb23@buffalo.edu>
%% 
%% This file may be distributed and/or modified under the conditions of
%% the LaTeX Project Public License, either version 1.2 of this license
%% or (at your option) any later version.  The latest version of this
%% license is in:
%% 
%%    http://www.latex-project.org/lppl.txt
%% 
%% and version 1.2 or later is part of all distributions of LaTeX version
%% 1999/12/01 or later.
%% 
\documentclass[a4paper,11pt,english,twoside]{ubthesis}
\renewcommand{\contentsname}{Table of Contents}
\usepackage{tocloft}
\renewcommand{\cftdotsep}{1}
\renewcommand{\cftpartleader}{\cftdotfill{\cftdotsep}} % for parts
\renewcommand{\cftchapleader}{\cftdotfill{\cftdotsep}} % for chapters
\renewcommand{\cftsecleader}{\cftdotfill{\cftdotsep}} % for sections,
\renewcommand{\cftsubsecleader}{\cftdotfill{\cftdotsep}}
\renewcommand\cftsecdotsep{\cftdot}
\renewcommand\cftsubsecdotsep{\cftdot}
\usepackage{setspace}
\doublespacing

%% You need to inclue this line or the hyperref package
\providecommand\phantomsection{}

\title{QUANTUM VIRIAL COEFFICIENTS VIA PATH INTEGRAL MONTE CARLO: THEORY AND DEVELOPMENT OF NOVEL ALGORITHMS}

\author{Ramachandran Subramanian}

\conferral{May 9, 2016}

\dept{Chemical and Biological Engineering}

\degree{phd}

\begin{document}
\begin{titlepage}
\maketitle
\end{titlepage}

\begin{ubfrontmatter}
\makecopyright
\cleardoublepage
\begin{acknowledgements}
    ``Mata, Pita, Guru, Deivam" is a saying in Sanskrit which translates to ``Mother, Father, Teacher, God" and it is the conventionally prescribed order in India, for expressing one's feeling of gratitude, appreciation and respect. Based on personal experiences, I intend to slightly deviate from this order.

    \noindent My mother (\emph{mummyji}): I wish to thank you for your unconditional love and relentless support throughout this endeavor. Your kindness and eagerness to help other people has always inspired me to become a better person.
    
    \noindent My father (\emph{subanna}): I wish to thank you for teaching and imparting me with strong moral values and discipline. Your constant criticism served as the perfect reality checks that I needed, when I was struggling for motivation.

    \noindent My sister (\emph{pete}): I wish to thank you for helping me make some of the most important decisions of my life and also for continuously believing in me.

    \noindent Dr. David A. Kofke: Words cannot express the amount of gratitude, respect and admiration that I have for you but still, seeing as I'm breaking rules anyway, I shall make an attempt. I wish to thank you for offering me this extraordinary opportunity to learn, interact and perform research with some of the best minds in the world. I am truly grateful to you for giving me the opportunity and support to present our work at various conferences. Your generosity, kindness and ubiquitous smile always made me feel welcome whenever I approached you for any advice. Thanks to you, I feel that I am better at communicating my ideas and thoughts more clearly. Thank you also for inspiring and instilling in me, a passion towards quantum mechanics, which ultimately led to my research focus. The amount of freedom that I experienced in the group and the sense of independence that it equipped me with, are deeply appreciated. The hospitality that you show your graduate students is of the highest quality and you always go above and beyond to make sure that we are comfortable. Thank you for making us feel home away from home. Last but definitely not the least, I am eternally indebted to you for making this a truly humbling and enlightening experience.

    \noindent Dr. Andrew J. Schultz: I wish to thank you for being patient with me all these years as I kept ``bugging" you with the most silly questions. Thank you for being the quintessential `Google' of the research group and always willing to help us out. 
\end{acknowledgements}
\tableofcontents
\cleardoublepage
\listoffigures
\cleardoublepage
\listoftables
\cleardoublepage
\begin{abstract}
This is my abstract
\end{abstract}
\end{ubfrontmatter}

\chapter{Introduction}

\begin{ubbackmatter}
\references[Bibliography]{}
\end{ubbackmatter}

\end{document}
\endinput
%%
%% End of file `thesis.tex'.
