% Place acknowledgments below.
``Mata, Pita, Guru, Deivam" is a saying in Sanskrit which translates to ``Mother, Father, Teacher, God" and it is the conventionally prescribed order in India, for expressing one's feeling of gratitude, appreciation and respect. Based on personal experiences however, I intend to slightly deviate from this order while acknowledging the people who have helped me become the man I am today.\\

\noindent My mother (\emph{mummyji}): I wish to thank you for your unconditional love and relentless support throughout this endeavor. Your kindness and eagerness to help other people has always inspired me to become a better person.\\

\noindent My father (\emph{subanna}): I wish to thank you for teaching and imparting me with strong moral values and discipline. Your constant criticism served as the perfect reality checks that I needed, when I was struggling for motivation.\\

\noindent Professor David A. Kofke: Words cannot express the amount of gratitude, respect and admiration that I have for you but still, seeing as I'm breaking rules anyway, I shall make an attempt. I feel extremely privileged to have worked under your guidance and also for being given this extraordinary opportunity to learn, interact and perform research with some of the best minds in the world. Your generosity, encouragement and ubiquitous smile always made me feel welcome whenever I approached you for any advice. Thanks to you, I feel that I am better at communicating my ideas and thoughts more clearly, a skill that means so much to me. Thank you also for inspiring and instilling in me, a passion towards quantum mechanics, which ultimately led to my research focus. I am truly grateful for your attitude towards students which provides for a stress-free environment in the lab. Your thought-provoking lectures, talks and presentations have always motivated me to become a better researcher. You are the most knowledgeable person that I know of, and just being in the presence of your wisdom for the last six years has had a profound impact in my professional as well as personal life. For that, I wish to thank you from the bottom of my heart. The hospitality and kindness that you show your students is of the highest quality as you always go above-and-beyond to make sure that we are comfortable. Thank you and your family for making us feel home away from home by means of group cookouts and barbecue events. Last but definitely not the least, I am eternally indebted to you for making this a truly humbling and enlightening experience.\\

\noindent Research Assistant Professor Andrew J. Schultz: I've always admired your passion for problem-solving and your commitment towards helping students. I've lost count of the number of times you have stayed late to help me understand something extremely fundamental, and for that, I tip my metaphorical hat to you. I am extremely grateful to you for introducing and helping me learn different software tools and applications. Your thoughtful suggestions and inputs have had a huge positive impact on the way I write code and have often saved me a world of suffering. I thank you also for taking time outside of your office hours and teaching me the right way to watch Star Wars. I'll always cherish our movie-marathon-sessions in the lab.\\

\noindent My committee member Professor Jeffrey R. Errington: I wish to thank you for your valuable comments and questions during group meetings, those have always motivated me to ask the right questions in research. I've often admired the confidence you exude while presenting in group meetings as well as conferences and I strive everyday to achieve it. I am sincerely grateful for the hospitality and kindness that you and your family have shown during numerous group events.  Thank you for serving on my committee.\\

\noindent My committee member Professor Johannes Hachmann: I cannot thank you enough for your excellent career advice and guidance. Your efforts to provide students with opportunities to connect and interact with people from the industry are sincerely appreciated. Thank you also for opening my eyes to Python, Data Science and many such interesting topics and presenting them in an awe-inspiring manner. I'd like to thank you for sharing your perspectives and experiences in graduate school which have helped me form my own, and also for serving on my committee.\\

\noindent Research collaborators: I would like to thank Dr. Allan H. Harvey(NIST) for insightful communications; Dr. Giovanni Garberoglio(Fondazione Bruno Kessler, Italy), Dr. Sergei Manzhos(University of Tokyo, Japan) and Dr. Massimiliano Bartolomei(IFF-CSIC, Spain) for helping me resolve issues regarding implementation of their computer codes.\\

\noindent My group members: I would like to thank Jung Ho Yang, for having a similar thought process as mine, which made developing a special bond that much easier; Weisong Lin, for motivating me to go to the gym more often; Pavan Behara, Shu Yang, Srihari Sankararaman, Saranya Harikrishnan, Nate Barlow, Katherine Shaul, Hye Min Kim, Nancy Cribbin, Tai Boon Tan, Sabry Moustafa and Chao Feng: for all their support, guidance and encouragement. I will surely treasure the memories I made whilst in your company.\\

\noindent My friends: Arjun Singh (\emph{phodu singh}), thank you for being my best friend and helping me through some of the toughest times I had ever faced despite going through some difficult times yourself. Dana Havas, thank you so much for being a really good friend whom I could trust and solicit personal and professional advice from. Vigneshwar Kumaran (\emph{jee}), thank you for accepting me just the way I am despite all my flaws. Thanks to all my friends at the CBE department(especially those from Prof. Errington's and Prof. Hachmann's research groups), badminton(especially Prof. Vipin Chaudhary, Biswa and Fahad), pool(especially Tariq Butt and Carl Galante), poker(especially Steve Zajac and John Putney), 33 Briarhurst gang and cricket(especially Anuj Atri and Amol Salunkhe).\\

\noindent My sister (\emph{pete}): I wish to thank you for helping me make some of the most important decisions of my life and also for continuously encouraging me no matter what.\\

\noindent I would like to thank the department's staff Dawn Townsend, Joan Wilson, Lori Duvall, Todd Nibbe and Marlo Kerr for being extremely helpful and ever so kind. Your efforts in making the lives of students much easier, are thoroughly and sincerely appreciated.\\

\noindent Finally, I'd like to thank all my relatives (especially Bala, Satyan, Harini and Abi) for their affectionate thoughts and company, and god for looking out for me.
